The \gaiaeso\ Survey: Combined estimates from multiple stellar analysis pipelines

### 2017-02-06 article/ms.tex
yolo

\newcommand{\acronym}[1]{{\small{#1}}}

\newcommand{\ges}{\acronym{GES}}
\newcommand{\project}[1]{\textsl{#1}}

\newcommand{\gaiaeso}{\project{Gaia-ESO}}

@@

\begin{document}
\title{The Gaia-ESO Survey: Combined estimates from multiple stellar analysis pipelines}

@@

\begin{abstract}
  Some stuff
\end{abstract}

@@

% Multiple analysis strategy: why

% primary advantage and disadvantage to multiple analysis strategy

% Description of two parameter rounds: stellar parameters, abundances

% What's reported by each node? Flags? Consistency of flags?

% Description of pipelines in appendix

% Commonalities between nodes: model atmospheres, line list, normalized spectra, initial params, spectrum library, xi calibration if necessary.

@@

% What do the nodes provide? Most parameters, flags, etc.

% CONTINUE from here:
% Stellar parameter determination

% Stellar parameter homogenisation

@@

    \item   \emph{The unbiased stellar parameters from any two nodes are correlated.}
            After accounting for individual node biases, we assume that the 
            stellar parameters reported by any two nodes will be correlated
            to some degree. The level of correlation between two nodes is unknown.
            This assumption arises from commonalities that led to the estimates from
            each node. Specifically: all nodes are instructed to use a common list 
            of line transitions, the same MARCS model photospheres, some nodes employ
            similar methods (e.g., spectral synthesis), and all nodes derive their 
            estimates from the same data (in some instances, even the same 
            continuum-normalized spectra are used).

            We note that even two nodes with identical methods are not guaranteed
            to be perfectly correlated, because the \emph{choice} of which transitions
            or spectral regions to use is not prescribed.

    \item   \emph{There are gaps in the data.} 
            No spectrum is guaranteed to have stellar parameters reported by all
            nodes. This assumption is a consequence of the different expertise in
            \gaiaeso: no node or method is expert across the entire range of parameter
            space explored. If stellar parameters are not reported, then the reason 
            must be specified through a \texttt{TECH} flag (see Section \ref{sec:flags}
            on flag usage).

@@

    \item   \emph{All `well-studied' stars have noisy measurement, and their values
            are not known with infinite precision.} But we assume they are unbiased!

@@

>>>>

\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\article}{\textit{Article}}

\newcommand{\ges}{\acronym{GES}}
\newcommand{\gaiaeso}{\project{Gaia-ESO}}

@@

\begin{document}
\title{The \gaiaeso\ Survey: Combined estimates from multiple stellar analysis pipelines}

@@

\begin{abstract}
	% The GES scientific goals
	% GES has a unique analysis strategy.
	% Allows XYZ benefits.
	% Here we describe the process for homogenisation of stellar parameters for FGK-type
	% stars analysed with both UVES and GIRAFFE data, in order to produce recommended results.
	% We construct a noise model for each node, and infer error behaviour for each node as a
	% function of S/N and stellar parameters.
	% We combine these to produce results.
	% For the first time, we describe blind tests conducted in GES to examine subjectivity, etc.
	% The typical reported uncertainties in stellar parameters are XYZ
	% We outline a list of recommended best practices and lessons learned for stellar parameter determination
	% We provide a common, homogeneous metallicity scale for GCs and OCs, and demonstrate that GES results can be used to infer diffusion/settling in clusters??
	% Other potential science highlights:
	% --> Something with velocity/metallicity in the field? two-point correlation function.
	% --> High-/low-alpha sequence with GIRAFFE stuff.

\end{abstract}

@@

The unique parallel analysis strategy adopted by the \ges\ provides a number of
benefits over other spectroscopic surveys.  We can identify the strengths and
weaknesses of state-of-the-art analysis methods, quantify the systematic 
differences that arise from \emph{defensible analysis decisions}, and produce an 
unbiased estimator of stellar parameters with representative uncertainties that account
for both random and systematic contributions.  Although this procedure requires an
enormous amount of collaboration and coordination, it is necessary to provide
a thorough evaluation of systematic uncertainties present between analysis 
pipelines that are routinely used in the literature.

We perform the \ges\ spectroscopic analysis for each data release in two 
consecutive rounds.  In the first round, the stellar parameters are reported by
each analysis group.  The homogenisation of stellar parameters is performed by the
working group leads (WG10, WG11, WG12, WG13), and the recommended results from each
working group are submitted to the top-level homogenisation group, WG15. WG15 makes
checks to ensure the results from different working groups are on the same scale, 
and combines the results from multiple working groups to produce a single set of 
\emph{Survey}-level recommended stellar parameters for each star \citep{Hourihane:2017}.
These recommended results are used in the second analysis round, where individual 
nodes report detailed chemical abundances given the recommended stellar parameters.
In this \article\ we describe only the first round of analysis; the homogenisation 
of detailed chemical abundances for FGK-type stars will be described in a companion 
contribution (Casey et al., 2017).

\subsection{Analysis methods}
\label{sec:analysis-nodes}
% How many nodes contributed to DR5 in WG11/WG10, and how many results did they provide per setup? (what fraction?)

% Summary description of nodes
% What do the nodes provide? Most parameters, flags, etc.
% Detailed description of pipelines in appendix

@@

\subsubsection{Summary of (dis-)similarities in analysis methods}
\label{sec:analysis-node-similarities}

Despite the multiple analysis strategies adopted in \ges, there are a number of
commonalities between the nodes. Specifically, all FGK-analysis nodes use the same:
\begin{enumerate}
  \item Updated MARCS model photospheres \citep{BG}.
  \item Atomic and molecular line lists, updated with recently measured oscillator strengths \citep{Heiter:2016,someone}.
  \item Rest-frame, pseudo-continuum-normalised spectra are provided to all nodes by the data reduction group. Individual echelle orders are also provided. Inverse variance arrays are available for all flux arrays.
  \item Initial stellar parameters are provided for all stars observed with GIRAFFE.
  \item Infrared photometry is available for all targets.
  \item For methods that perform template-fitting, all nodes are provided with a grid of high-resolution continuum-normalised spectra synthesised using (1) and (2).  The synthetic grid covers the entire wavelength region used for \ges.
  \item Where relevant, nodes are provided with empirical microturbulence calibrations for main-sequence, sub-giant, and giant stars. These relations are given in Appendix \ref{app:xi}.
\end{enumerate}

There are three caveats to the listed items above.  Firstly, while nodes are encouraged
to use the same empirically-scaled interpolator for photospheric models \citep{Masseron:2006},
this is not required.  Secondly, the nodes are not required to use the 
pseudo-continuum-normalised spectra provided to them; each node can chose to 
adopt their own continuum normalisation procedure.  Lastly, although all nodes use the 
same atomic and molecular data, the nodes are free to chose \emph{which} lines (or 
spectral regions) to use when fitting their models to the data.

In summary, the analysis nodes generally differ only in \emph{defensible analysis decisions}:
there is no well-founded \emph{a priori} reason that would explain why one method should be
significantly better or worse across the space of stellar parameters.  For these reasons,
the Gaia-ESO Survey is well-placed to critically examine what analysis decisions will
objectively result in more accurate or precise results.

@@

    \item   \emph{The unbiased stellar parameters from any two nodes are correlated.}
            After accounting for individual node biases, we assume that the 
            stellar parameters reported by any two nodes will be correlated
            to some degree. The level of correlation between two nodes is unknown.
            This assumption arises from commonalities that led to the estimates from
            each node (Section \ref{sec:analysis-node-similarities}).

            We reiterate that even two nodes with identical methods are not guaranteed
            to be perfectly correlated, because the \emph{choice} of which transitions
            or spectral regions to use is not prescribed.

    \item   \emph{There are gaps in the data.} 
            No spectrum is guaranteed to have stellar parameters reported by all
            nodes. This assumption is a consequence of the different expertise in
            \gaiaeso: no node or method is expert across the entire range of parameter
            space explored.

@@

    \item   \emph{All `well-studied' stars have noisy measurement, and their values
            are not known with infinite precision.} That is to say, we assume that there
            are some well-studied stars with unbiased measurements, but those
            measurements are noisy. For this reason, we must include the stellar
            parameters of each well-studied star as a model parameter, and use 
            non-spectroscopic data to place a strong prior on that stellar parameter.

@@

\newpage
\appendix
\section{Microturbulence calibrations} \label{app:xi}

% Version 2.0 taken from http://great.ast.cam.ac.uk/GESwiki/GesWg/GesWg11/Microturbulence

The following microturbulence calibrations were recommended for nodes to use where appropriate. For main-sequence or sub-giant stars with $\teff >= 5000$~K and $\logg >= 3.5$,

\begin{equation}
\xi = 1.05 + 2.51\times10^{-4}(\teff T_0) + 1.5\times10^{-7}(\teff - T_0)^2 - 0.14(\logg - g_0) - 0.005(\logg - g_0)^2 + 0.05\feh + 0.01\feh^2 \quad ,
\end{equation}

\noindent{}and for main-sequence or sub-giant stars with $\teff < 5000$~K,

\begin{equation}
\xi = 1.05 + 2.51\times10^{-4}(5000 - T_0) + 1.5\times10^{-7}(5000 - T_0)^2 - 0.14(\logg - g_0) - 0.005(\logg - g_0)^2 + 0.05\feh + 0.01\feh^2 \quad ,
\end{equation}

\noindent{}and for giant stars:

\begin{equation}
\xi = 1.25 + 4.01\times10^{-4}(\teff - T_0) + 3.1\times10^{-7}(\teff - T_0)^2 - 0.14(\logg - g_0) - 0.05(\logg - g_0)^2 + 0.05\feh + 0.01\feh^2 \quad .
\end{equation}

\noindent{}In all three cases above: $T_0 = 5500$~K and $g_0 = 4.0$.

\section{Flags} \label{app:flags}

\section{\texttt{Stan} code for homogenisation noise model} \label{app:stan-code}
